Lattice Boltzmann study of mass transfer for complicated geometries

### 2012-01-27 Text/introduction.tex
Minor changes in text

\section{Mass transfer definitions}
By the definition the mass transfer coefficient from the surface with the imposed constant
concentration $\cbubble$ is the following:
\beq
k_L=\frac{\dot{m}}{P \Delta C},
\feq
where $\dot{m}$ is the mass rate $\Bigl[\frac{kg}{s}\Bigr]$, $P$ is the area of the surface
$\Bigl[m^2\Bigr]$, $\Delta C$ is the concentration difference between the surface and the medium
$\Bigl[\frac{kg}{m^3}\Bigr]$. Therefore, $k_L$ has a dimension of velocity
$\Bigl[\frac{m}{s}\Bigr]$. Usually, the medium concentration is taken at the infinite distance
from the bubble. However, in the case of complicated geometries and inhomogenuous concentrations, 
the medium concentration can be the average concentration in the domain, the flux averaged
concentration at the outlet. Thus, one needs to establish a thorough definition of the volumetric
mass transfer coefficient in the case of complex geometries and hydrodynamic velocity patterns.

There are different methods to estimate the mass transfer coefficient $k_L$. We first examine the
theoretical definitions of the mass transfer in case of the point sources.
\subsection{Point mass sources}
In what follows we will present three approaches to calculate point mass transfer
coefficients:
\begin{enumerate}
\item
Let us look at the infinitisemal small domain of the volume $A \Delta x$ in a steady-state, not
moving and with the mass source. Then the concentration difference can be found as $\Delta C = C^* -
C(t)$, where $C^*$ is the imposed point source concentration, $C(t)$ is
the time depending average concentration. Therefore one can write the time dependent ODE for the
average concentration in the domain:
\beq
\dot{m}= A \Delta x \frac{\mathrm{d}C}{\mathrm{d} t} = k_L P (C^{*}-C(t)), 
\feq
with the initial condition $C(0)=0$
The solution in time becomes as follows:
\beq
C(t)= C^{*}(1-\exp(-\vol t )), 
\feq
where $\vol$ is the volumetric mass transfer coefficient defined as:
\beq
\vol=k_L \frac{P}{V},
\feq
where $P$ is the source surface, $V$ is the unit cell volume.
\item
The previous consideration is able to predict mass transfer happening in moving with the velocity
$U$ liquid, see Fig. \ref{fig:moving:frame}.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/mass_transfer.eps}
\caption{The mass transfer in the moving liquid. \label{fig:moving:frame}}
\end{figure}

The mass accumulated in the unit volume $V=A \Delta x$ can be calculated as the difference of 
mass fluxes entering and leaving domain $U \bigl(C(x+\Delta x)-C(x)\bigr)$. We assume that the
source is a point source with the effective perimeter P. The accumulated mass should be proportional
to the mass transfer coefficient:
\beq
U \bigl(C(x+\Delta x)-C(x)\bigr)=k_L P \bigl(C^{*}-C(x)\bigr), 
\feq 
giving the same equation but only in the coordinate domain:
\beq
C(x)= C^{*}\Bigl(1-\exp\bigl(-k_L a \frac{x}{U} \bigr)\Bigr).
\label{main:mass:transfer:expression} 
\feq

>>>>

\section{Mass transfer definitions}
By the definition the mass transfer coefficient from the surface with the imposed constant
concentration $\cbubble$ is the following:
\beq
k_L=\frac{\dot{m}}{P \Delta C},
\feq
where $\dot{m}$ is the mass rate $\Bigl[\frac{kg}{s}\Bigr]$, $P$ is the area of the surface
$\Bigl[m^2\Bigr]$, $\Delta C$ is the concentration difference between the surface and the surrounding medium
$\Bigl[\frac{kg}{m^3}\Bigr]$. Therefore, $k_L$ has a dimension of velocity
$\Bigl[\frac{m}{s}\Bigr]$. Usually, the surrounding medium concentration is taken at the infinite distance
from the bubble. However, in the case of complicated geometries and non-homogeneous concentrations, 
the medium concentration can be the average concentration in the domain, the flux averaged
concentration at the inlet or outlet. Thus, one needs to establish a thorough definition of the volumetric
mass transfer coefficient in the case of complex geometries and non-trivial hydrodynamic velocity patterns.

There are different methods to estimate the mass transfer coefficient $k_L$. We first examine the
theoretical definitions of the mass transfer in case of point mass sources.
\subsection{Point mass sources}
In what follows we will present three approaches to calculate point mass transfer
coefficients:
\begin{enumerate}
\item
Let us look at the infinitesimal small domain of the volume $A \Delta x$ not
moving and with the point mass source. Then the concentration difference can be found as $\Delta C = C^* -
C(t)$, where $C^*$ is the imposed point source concentration, $C(t)$ is
the time dependent average concentration. Therefore one can write the time dependent ODE for the
average concentration in the domain:
\beq
\dot{m}= A \Delta x \frac{\mathrm{d}C}{\mathrm{d} t} = k_L P (C^{*}-C(t)), 
\feq
with the initial condition $C(0)=0$
The solution can be found by solving ODE:
\beq
C(t)= C^{*}(1-\exp(-\vol t )), 
\feq
where $\vol$ is the volumetric mass transfer coefficient defined as:
\beq
\vol=k_L \frac{P}{V},
\feq
where $P$ is the source surface, $V$ is the unit cell volume.
\item
The previous consideration is able to predict mass transfer happening in moving with the velocity
$U$ liquid, see Fig. \ref{fig:moving:frame}. 
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/mass_transfer.eps}
\caption{The mass transfer in the moving liquid. \label{fig:moving:frame}}
\end{figure}

If one can assume that the mass point sources are distributed in the whole medium, then the mass accumulated in the unit volume $V=A \Delta x$ can be calculated as the difference of 
mass fluxes entering and leaving domain $U \bigl(C(x+\Delta x)-C(x)\bigr)$. We assume that the
source is a point source with the effective perimeter (area) $P$. The accumulated mass should be proportional
to the mass transfer coefficient:
\beq
U \bigl(C(x+\Delta x)-C(x)\bigr)=k_L P \bigl(C^{*}-C(x)\bigr), 
\feq 
giving the same equation but only in the coordinate domain:
\beq
C(x)= C^{*}\Bigl(1-\exp\bigl(-k_L a \frac{x}{U} \bigr)\Bigr).
\label{main:mass:transfer:expression} 
\feq
